### 2016-11-17 chapter1.tex
Summary of observations of asymmetry

There have been at least 6 mechanisms proposed for how polar patches form on the dayside.
\begin{enumerate}
	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure
	\item Average flow patterns varying in time \citep{Anderson1988}
	\item Transient magnetopause reconnection \citep{Lockwook1992}
	\item Plasma production by cusp particle precipitation \citep{Roger1994,Millward1999}
	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
	\item Alfven wave coupling \citep{Prikryl1999}
\end{enumerate}
Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Roger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.  These high densities must originate from the resivor of solar illuminated plasma on the dayside.  Additionally, a mechanism is still required to separate patches from the cusp region.  High speed flow channels can "cut" segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.  Models have also shown changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.  However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.

@@

@@

-Patch formation here?
-Dahlgren 2012 a, b
-Semeter 2009
-Lyne 2015 (rockets)
-Moen 2012

>>>>

There have been at least 6 mechanisms proposed for how polar patches form on the dayside.
\begin{enumerate}
	\item Discrete changes in solar wind parameters, usch as IMF, density, speed, and pressure \citep{Sojka1994}
	\item Average flow patterns varying in time \citep{Anderson1988}
	\item Transient magnetopause reconnection \citep{Lockwook1992}
	\item Plasma production by cusp particle precipitation \citep{Roger1994,Millward1999}
	\item Plasma flow jet channels that cut continuous streaches of plasma into segments 	\citep{Valladares1998}
	\item Alfven wave coupling \citep{Prikryl1999}
\end{enumerate}
Although particle precipitation has been shown to create enhanced plasma density in the cusp region \citep{Roger1994}, it is unreasonable for this mechanism to account for the largest density enhancements observed in polar patches.  These high densities must originate from the resivor of solar illuminated plasma on the dayside.  Additionally, a mechanism is still required to separate patches from the cusp region.  High speed flow channels can "cut" segments off of a tongue of high-density plasma that has been pulled into the polar cap by convection \citep{Valladares1994,Valladares1998}.  Models have also shown changing large scale convection patterns quickly in time does produce density structures similar to patches \citep{Anderson1988}.  However, the mechanism that seems to be dominant for the majority of polar patches observed is transient magnetic reconnection \citep{Carlson2012}.

@@

\citet{Coley1998} examined the occurrence of polar patches in both the northern and southern hemispheres with DE-2 and found that the frequency of patch occurrence was greatest when  the cusp was slightly dayward of the terminator.  This creates the situation of an unsunlit polar cap, so background plasma density is low and blobs of enhanced density are easily visible.  In addition, the cusp then provides a gateway of highly ionized sunlit plasma into the dark polar cap.  \citet{Rodger1996} also considered occurrence of patches but with a HF radar in Halley, Antarctica.  The diurnal variation in patch occurrence was found to peak at magnetic noon and the seasonal variation in equinox months.  The only correlation between hourly averaged solar wind parameters that was found was a bias towards patches occurring with a negative IMF \(B_z\) component.

@@

Recent advancements have improved the ability to image density structures in the ionosphere, particularly using multi-instrument approaches.  \citet{Semeter2009} introduced a method by which the density pattern in a particular volume could be imaged in three dimensions using AMISR systems.  This technique was later used by \citet{Dahlgren2012a,Dahlgren2012b} to image a polar patch.  \citet{Dahlgren2012b} additionally used ASIPs and SuperDARN radars to investigate the polar patch and found that backscatter tended to be observed more on the trailing edge of the patch.  This agrees well with many studies that have examined the asymmetry of small-scale plasma structuring surrounding plasma patches.  This asymmetry was first observed by \citet{Weber1984}.  Later, \citet{Milan2002b} found that backscatter power from HF radars tended to be greater on the trailing edges of moving structures, which was attributed to GDI being unstable on the trailing edge but stable on the leading edge.  \citet{Koustov2012} found similar results examining sun-aligned arcs.  HF radar backscatter power tended to be greater trailing behind a sun-aligned arc moving through the polar cap that was identified with ASIP data.  \citet{Moen2012} made direct measurements of plasma density structuring using the ICI-2 sounding rocket.  This study provided evidence that decameter scale plasma structuring (the same as observed by HF radars) had spawned from kilometer scale density gradients and structuring was greatest where GDI was operational.  \citet{Hosokawa2016} also considered the asymmetry of patches using an all-sky airglow imager (ASI) and found that the gradient on the leading edge of patches was typically 2--3 times steeper than that on the trailing edge.  In addition, they found large finger-like structures on the trailing edge of patches that were 10s--100s of kilometers in size.  These types of structures agree with GDI predictions, in particular they match the simulations of \citet{Gondarenko2004}.
